A White Noise Approach to Evolutionary Ecology

### 2020-04-10 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
improved proof of stable equilibrium for logistic growth and stabilizing selection

where $\bar x(t)$ and $\sigma^2(t)$ are respectively the mean trait and phenotypic variance at time $t\geq0$. Second, we assume selection is determined by the growth rate $m(h,x)$ that is differentiable with respect to both arguments and is bounded above by some $r\in\mathbb{R}$ for all $x\in\mathbb{R}$ and $h\geq0$. We denote by $K$ an operator that takes functions such as $\nu(x,t)$ as its argument and returns other functions that may still depend on the spatial variable $x$ and the temporal variable $t$. For our application we will be focused on operators $K$ that can be written as $(K\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x,y)\nu(y,t)dy$ for some non-negative and bounded bivariate function $\kappa$. We set $h=(K\nu)(x,t)$ to account for nonlocal effects of the abundance density $\nu(x,t)$ on the fitness of individuals with trait value $x$ [@Volpert2014]. An example of a nonlocal effect comes from competition for a common resource among all individuals in a population regardless of their trait value. In this case all individuals in the population effect each others fitness' even though they may exhibit diverse trait values. The term nonlocal comes from interpreting the trait value as spatial location. From here on we abbreviate $m(h,x)=m\big((K\nu)(x,t),x\big)$ to just $m(\nu,x)$. Third, we assume mutation is captured by diffusion with coefficient $\frac{\mu}{2}$. Now that these technicalities have been treated we can model the demographic dynamics of a population and the dynamics of a quantitative character simultaneously by the PDE

@@

with the initial condition $\nu(x,0)$ described above. This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@lawrenceevans2010]. Interpreting the trait value $x$ as spatial location and ignoring nonlocal effects, which can be satisfied by setting $K\nu=\nu$, this model of spatially distributed population dynamics has been intensely studied [for a review, see @Cantrell2004]. When $\mu=0$, equation (\ref{eq1}) can be seen as an analog of equation (\ref{first}) for a continuum of types. By assuming mutation acts as diffusion, the effect of mutation causes $\nu(x,t)$ to flatten out over time. In fact, if the growth rate is constant across $x$, then this model of mutation will cause $\nu(x,t)$ to converge to a flat line as $t\to\infty$. Although clearly an idealized representation of biological reality, this model is sufficiently general to capture a large class of dynamics including density dependent growth and frequency dependent selection. As an example, logistic growth combined with quadratic stabilizing selection can be captured using the growth rate 

@@

where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x,y)=1$. This exemplary fitness function has a few convenient properties that we breifly take note of. First, the effect of competition induces a local carrying capacity on the population, leading to a finite equilibrium abundance over bounded subsets of $\mathbb{R}$. Second, abiotic selection prevents the abundance density from diffusing too far from the abiotic optimum, leading to a finite equilibrial abundance across all of $\mathbb{R}$, $\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, finite equilibrial mean trait $\lim_{t\to\infty}\bar x(t)=\theta$ and finite equilibrial phenotypic variance $\lim_{t\to\infty}\sigma^2(t)=\sqrt{\mu/a}$, so long as $\bar x(0)\in\mathbb{R}$, $\sigma^2(0),N(0)\in(0,+\infty)$ and $r>\sqrt{a\mu}/2$. We prove these statements in SM \S\ref{equilib}. By replacing $\kappa(x,y)=1$ with a particular Gaussian function, we will find in \S\ref{coev} and SM \S\ref{diffuse} this fitness function becomes equivalent to a fitness function derived from niche theory that we use to model coevolution driven by resource competition.

@@

\begin{multline}\label{p_PDE}
\frac{\partial}{\partial t}p(x,t)=\frac{\partial}{\partial t}\frac{\nu(x,t)}{N(t)}=\frac{1}{N^2(t)}\left(N(t)\frac{\partial}{\partial t}\nu(x,t)-\nu(x,t)\frac{d}{dt}N(t)\right) \\
=\frac{1}{N(t)}\Big(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)-\bar m(t)\nu(x,t)\Big) \\
=(m(\nu,x)-\bar m(t)) p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t).
\end{multline}
This result closely resembles Kimura's continuum-of-alleles model [@Kimura1965; @Brger1986]. The primary difference being that our model utilizes diffusion instead of convolution with an arbitrary mutation kernel. Of course, our model of mutation can be derived as an approximation to Kimura's model, which has been referred to as the Gaussian allelic approximation in reference to the distribution of mutational effects at loci in a genome on the values of traits [@Lande1975; but see also @9780471986539], the infinitesimal model in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect [@Barton2017] and the Gaussian descendants approximation in reference to offspring trait values being normally distributed around their parental values [@Turelli2017]. Alternatively, since diffusion is the continuous-time equivalent to convolution against a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can also be seen as a special case of Kimura's continuum-of-alleles model. We return to these concepts in \S\ref{inheritance} where we review a simple model of inheritance that leads us to an expression for the stochastic evolution of additive genetic variance.

@@

Recall, we use $\nu(x,t)$ to denote the density of a superprocess, given it exists. Assuming the interactions manifest only in the reproductive law and that spatial movement follows Brownian motion with diffusion parameter $\sqrt\mu$ independent of both $X_t$ and $x$, Li [-@zeng1998absolute] proved a result that implies the interacting superprocess on one dimensional trait space has a density $\nu(x,t)$ which is non-negative, integrable, continuous in time and space and satisfies the SPDE
\begin{equation}\label{SPDE}
\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
\end{equation}
Comparing equation (\ref{SPDE}) to equation (3.5) of Li [-@zeng1998absolute], our $m$ and $V$ correspond to Li's $b$ and $c$ respectively. It is important to note that, under the assumptions made in Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] and Li [-@zeng1998absolute], equation (\ref{SPDE}) is only formal when $m(h,x)$ is bounded across all combinations of $h\geq0$ and $x\in\mathbb{R}$. However, recalling our condition $m(h,x)\leq r\in\mathbb{R}$, the growth rates we consider are only bounded above. Yet, in the proof of the construction of the interacting superprocess as the limit of rescaled branching diffusions, Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] assumed $m(h,x)$ to be bounded to guarantee the total mass process will have finite mean and variance, for finite $t\geq0$. This allowed the authors to employ a tightness criterion for sequences of measures and show the rescaled processes converge to a superprocess with finite total mass. Li's [-@zeng1998absolute] result builds directly off of Méléard and Roelly's construction, inheriting the assumption of boundedness for $m(h,x)$. However, in Li [-@zeng1998absolute], the sufficiency of $m(h,x)$ being bounded above is even more clear since Li works explicitly with a common upperbound for both $m(h,x)$ and $V$. Hence, one can repeat the necessary proofs replacing the assumption that $m(h,x)$ is bounded with the assumption that $m(h,x)$ is merely bounded above to derive the same results.

@@

Our treatment of the relationship between breeding values and expressed traits follows classical quantitative genetic assumptions such as those used in the seminal paper by Lande [-@Lande1975] to investigate the maintenance of genetic variation. In particular, we assume that the expressed trait for any given individual is independent of environmental conditions and normally distributed around their breeding value with variance $\eta$. Hence, $\sigma^2=G+\eta$. In the case that all of the effects of alleles on an expressed trait are additive, $\eta$ is known as the _variance of environmental deviation_ [@Lande1975; @michaellynch1998]. For a given breeding value, we denote the probability density of a randomly drawn expressed trait by $\psi(x,g)$. Hence,
\begin{equation}
\psi(x,g)=\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g)^2}{2\eta}\right).
\end{equation}
To include this relationship in our framework, we write $\rho(g,t)$ as the abundance density of breeding values at time $t$ so that $\int_{-\infty}^{+\infty}\rho(g,t)dg=\int_{-\infty}^{+\infty}\nu(x,t)dx=N(t)$. We switch our focus from directly modelling the evolution of $\nu(x,t)$ to modelling the evolution of $\rho(g,t)$. Once $\rho(g,t)$ is determined, we can compute $\nu(x,t)$ via
\begin{equation}\label{nu_from_rho}
\nu(x,t)=\int_{-\infty}^{+\infty}\rho(g,t)\psi(x,g)dg.
\end{equation}
However, since selection acts on expressed phenotypes, we use the assumed relationship between breeding values and expressed traits to calculate the fitness of breeding values. To motivate the approach taken, consider the problem of inferring the breeding value of an individual given its expressed trait $x$. Denote $\mathfrak{g}$ a random variable representing the unkown breeding value. Under this model of inheritance we know $x$ is a random sample from a normal distribution with mean $\mathfrak{g}$ and variance $\eta$. Maximizing likelihood suggests $x$ is our best guess for $\mathfrak{g}$, but the actual value of $\mathfrak{g}$ is normally distributed around $x$ with the variance $\eta$. Hence, for fixed $x$, we obtain $\psi(x,g)$ as the probability density of $\mathfrak{g}$. Thus, the mean fitness of a breeding value $g$ across all individuals carrying $g$ can be written as
\begin{equation}\label{m_rel}
m^*(\rho,g)=\int_{-\infty}^{+\infty}m(\nu,x)\psi(x,g)dx.
\end{equation}
This is similar to the approach taken by Kimura and Crow [-@Kimura1978] to calculate the overall effects of selection for expressed characters onto the changes in the distribution of alleles encoding those characters. However, instead of focusing on the frequencies of alleles at particular loci, our results focus on the densities of breeding values. With the relationship between $m(\nu,x)$ and $m^*(\rho,g)$ established, we define the evolution of $\rho(g,t)$ by the SPDE
\begin{equation}\label{rho_SPDE}
\dot\rho(g,t)=m^*(\rho,g)\rho(g,t)+\frac{\mu}{2}\frac{\partial^2}{\partial^2 g}\rho(g,t)+\sqrt{V\rho(g,t)}\dot W(g,t).
\end{equation}
Equation (\ref{rho_SPDE}) is a stochastic generalization of the deterministic PDE (\ref{eq1}) from \S\ref{deterministic}, but describes the evolution of the distribution of breeding values instead of expressed characters. In \S\ref{wnc} we review the origins of this equation and provide some theory to help make sense of it, particularly the term $\dot W$.

@@

```{r, message=FALSE, echo=F, fig.align='center', out.width='100%', fig.cap='\\label{temporal}Temporal dynamics of mean trait (top), additive genetic variance (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right). Red lines indicate average trend across species.'}
knitr::include_graphics('community_dynamics.jpg')
```

@@

We now make use of the expressions derived for competition coeffecients and selection gradients to investigate their relationship. As a first pass, let us assume all model parameters are equivalent across species and that each species has the same abundance and trait variance. Let us further assume that richness $S$ is large and the distribution of mean trait values is normal with mean $\bar{\bar x}$, variance $V_{\bar X}$ and density $f_{\bar X}$. Such assumptions are typical when deriving analytical results in the field of theoretical coevolutionary community ecology [@GuimaresJr2011; @Guimares2017; @Nuismer2012; @Nuismer2018]. If $\bar{\bar x}$ is near $\theta$ and $V_{\bar X}$ is much smaller than $|2R/a-G-\eta|$, then we may approximate $r_i$ with

@@

We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise, demonstrated our approach by deriving a model of diffuse coevolution and investigated the relationship between selection gradients and competition coefficients, finding that these quantities exhibit interesting relationships which shed light on the feedback between the structure and dynamics of ecological communities.

Our approach has the merit of rigorously synthesizing the dynamics of abundance and phenotypic evolution for populations experiencing demographic stochasticity. Yet, there remains biological details and their associated technical challenges that need to be confronted for gaining a more thorough and rigorous understanding of ecological communities. We touch on just four of them here and provide suggestions for future research to approach these challenges.

__Limitations of diffusion limits__

As noted by Feller [-@feller1951], diffusion limits provide reasonable approximations for large populations, but relatively small populations require discrete models. Hence, as a diffusion limit, SPDE (\ref{SPDE}) restricts parameter values to regions that maintain large population sizes. Although this suggests an important restriction on any model developed under this framework by implying populations are not at risk of extinction, the SDE describing abundance dynamics technically permits extinction. However, for small abundances, pathology emerges in the evolution of trait means and variances. In particular, stochastic components of the SDE describing the evolution of $\bar x$ and $\sigma^2$ diverge towards infinite values as $N\to0$. Hence, studies of evolutionary rescue and colonization-extinction dynamics that incorporate phenotypic evolution should be pursued via a different approach. A natural alternative can be developed utilizing the underlying BBM that SPDE (\ref{SPDE}) is a diffusion limit of (see section \ref{stochastic}). This process explicitly tracks individuals throughout their life-history and captures reproduction as branching events. Hence, BBM processes model population size as a non-negative integer instead of a continuously varying number. In particular, the pathological behavior described for the diffusion limit does not occur for BBM as population size goes to zero.

@@

Here we set out to show, given appropriate initial conditions, $N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, $\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta$ and $\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}$ when
\begin{equation}\label{m_log_stab}
m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(y,t)dy
\end{equation}
for $\theta\in\mathbb{R}$, $\mu\geq0$, $a,c>0$ and $r>\tfrac{1}{2}\sqrt{\mu a}$. When it exists, set $\nu_\infty(x)=\lim_{t\to\infty}\nu(x,t)$. The mean fitness can be calculated as
\begin{equation}
\bar m(t)=r-\frac{a}{2}((\theta-\bar x(t))^2-\sigma^2(t))-cN(t), \ \ \forall \ \ t\geq0.
\end{equation}
Solving for $N_\infty$ amounts to setting $\dot N(t)=0$ and solving for $N(t)$. This reduces to solving $\bar m(t)=0$ for $N(t)$, which returns
\begin{equation}
N_\infty=\frac{1}{c}\left(r-\frac{a}{2}\left((\theta-\bar x_\infty)^2-\sigma^2_\infty\right)\right).
\end{equation}
To find $\nu_\infty(x)$ we solve $\tfrac{\partial}{\partial t}\nu(x,t)=0$ for $\nu(x,t)$. This implies the following ordinary differential equation
\begin{equation}
\frac{d^2}{dx^2}\nu_\infty(x)=\left(\frac{2c}{\mu}N_\infty+\frac{a}{\mu}(\theta-x)^2-\frac{2r}{\mu}\right)\nu_\infty(x)
\end{equation}
which has the solution
\begin{equation}
\nu_\infty(x)=\frac{N_\infty}{\sqrt{2\pi}}\left(\frac{a}{\mu}\right)^{\frac{1}{4}}\exp\left(-\frac{(\theta-x)^2}{2\sqrt{\mu/a}}\right).
\end{equation}
From this expression we infer $\bar x_\infty=\theta$ and $\sigma^2_\infty=\sqrt{\mu/a}$. Hence $N_\infty=\frac{1}{c}\left(r-\frac{1}{2}\sqrt{a\mu}\right)$. This holds for any initial $\nu(\cdot,0)\in L^1(\mathbb{R})$ such that $\bar x(0)\in\mathbb{R}, \ \sigma^2(0),N(0)\in(0,+\infty)$.

>>>>

where $\bar x(t)$ and $\sigma^2(t)$ are respectively the mean trait and phenotypic variance at time $t\geq0$. Second, we assume selection is determined by the growth rate $m(h,x)$ that is differentiable with respect to both arguments and is bounded above by some $r\in\mathbb{R}$ for all $x\in\mathbb{R}$ and $h\geq0$. We denote by $K$ an operator that takes functions such as $\nu(x,t)$ as its argument and returns other functions that may still depend on the spatial variable $x$ and the temporal variable $t$. For our application we will be focused on operators $K$ that can be written as $(K\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x,y)\nu(y,t)dy$ for some non-negative and bounded bivariate function $\kappa$. We set $h=(K\nu)(x,t)$ to account for nonlocal effects of the abundance density $\nu(x,t)$ on the fitness of individuals with trait value $x$ [@Volpert2014]. An example of a nonlocal effect comes from competition for a common resource among all individuals in a population regardless of their trait value. In this case all individuals in the population effect each others fitness' even though they may exhibit diverse trait values. The term nonlocal comes from interpreting the trait value as spatial location. From here on we abbreviate $m(h,x)=m\big((K\nu)(x,t),x\big)$ to just $m(\nu,x)$. Third, we assume mutation is captured by diffusion with coefficient $\frac{\mu}{2}$. With these technicalities addressed we model the demographic dynamics of a population and the dynamics of a quantitative character simultaneously by the PDE

@@

with the initial condition $\nu(x,0)$ described above. This PDE is semilinear due to the dependency of the growth rate $m(\nu,x)$ on the solution $\nu(x,t)$ and is referred to as a scalar reaction-diffusion equation [@Henry1981; @lawrenceevans2010]. Interpreting the trait value $x$ as spatial location and ignoring nonlocal effects, which can be satisfied by setting $K\nu=\nu$, this model of spatially distributed population dynamics has been intensely studied [for a review, see @Cantrell2004]. When $\mu=0$, equation (\ref{eq1}) can be seen as an analog of equation (\ref{first}) for a continuum of types. By assuming mutation acts as diffusion, the effect of mutation causes $\nu(x,t)$ to flatten out over time. In fact, if the growth rate is constant across $x$, then this model of mutation will cause $\nu(x,t)$ to converge to a flat line as $t\to\infty$. Although clearly an idealized representation of biological reality, this model is sufficiently general to capture a large class of dynamics including density dependent growth and frequency dependent selection. As an example, logistic growth combined with quadratic stabilizing selection can be captured using the growth rate 
\begin{equation}\label{special_m}

@@

where $r\in\mathbb{R}$ is the intrinsic growth rate in the absence of abiotic stabilizing selection, $a>0$ is the strength of abiotic stabilizing selection around the phenotypic optimum $\theta\in\mathbb{R}$, $c>0$ captures the sensitivity of fitness to intraspecific competition and we have set $\kappa(x,y)=1$ so that competitive interactions cause the same reduction in fitness regardless of trait value. This exemplary fitness function has a few convenient properties that we breifly take note of. First, the effect of competition induces a local carrying capacity on the population, leading to a finite equilibrium abundance over bounded subsets of $\mathbb{R}$. Second, abiotic selection prevents the abundance density from diffusing too far from the abiotic optimum, leading to a finite equilibrial abundance across all of $\mathbb{R}$, $\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, finite equilibrial mean trait $\lim_{t\to\infty}\bar x(t)=\theta$ and finite equilibrial phenotypic variance $\lim_{t\to\infty}\sigma^2(t)=\sqrt{\frac{\mu}{a}}$, so long as $\bar x(0)\in\mathbb{R}$, $\sigma^2(0)\in[0,+\infty)$, $N(0)\in(0,+\infty)$ and $r>\frac{1}{2}\sqrt{a\mu}>0$. We prove these statements in SM \S\ref{equilib}. By replacing $\kappa(x,y)=1$ in expression (\ref{special_m}) with a particular Gaussian function, we will find in \S\ref{coev} and SM \S\ref{diffuse} this fitness function becomes equivalent to a fitness function derived from niche theory that we use to model coevolution driven by resource competition.

@@

\begin{multline}\label{p_PDE}
\frac{\partial}{\partial t}p(x,t)=\frac{\partial}{\partial t}\frac{\nu(x,t)}{N(t)}=\frac{1}{N^2(t)}\left(N(t)\frac{\partial}{\partial t}\nu(x,t)-\nu(x,t)\frac{d}{dt}N(t)\right) \\
=\frac{1}{N(t)}\Big(m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)-\bar m(t)\nu(x,t)\Big) \\
=(m(\nu,x)-\bar m(t)) p(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}p(x,t).
\end{multline}
This result closely resembles Kimura's continuum-of-alleles model [@Kimura1965; @Brger1986]. The primary difference being that our model utilizes diffusion instead of convolution with an arbitrary mutation kernel. Of course, our model of mutation can be derived as an approximation to Kimura's model, which, for sexually reproducing populations, has been referred to as the Gaussian allelic approximation in reference to the distribution of mutational effects at loci in a genome on the values of traits [@Lande1975; but see also @9780471986539], the infinitesimal model in reference to modelling continuous traits as being encoded by an infinite number of loci each having infinitesimal effect [@Barton2017] and the Gaussian descendants approximation in reference to offspring trait values being normally distributed around their parental values [@Turelli2017]. Alternatively, since diffusion is the continuous-time equivalent to convolution against a Gaussian kernel (SM \S\ref{diffconvequiv}), equation (\ref{p_PDE}) can also be seen as a special case of Kimura's continuum-of-alleles model. We return to these concepts in \S\ref{inheritance} where we review a simple model of inheritance that leads us to an expression for the stochastic evolution of additive genetic variance. To clarify the distinction between this model and previous models of phenotypic evolution we refer to it as the Deterministic Asexual Gaussian allelic model with Abundance dynamics (abbreviated DAGA). Later, we will extend this model to include the effects of demographic stochasticity, which we refer to as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA).

@@

Recall, we use $\nu(x,t)$ to denote the density of a superprocess, given it exists. Assuming the interactions manifest only in the reproductive law and that spatial movement follows Brownian motion with diffusion parameter $\sqrt\mu$ independent of both $X_t$ and $x$, Li [-@zeng1998absolute] proved a result that implies the interacting superprocess on one dimensional trait space has a density $\nu(x,t)$ which is non-negative, integrable, continuous in time and space and satisfies the SPDE
\begin{equation}\label{SPDE}
\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)+\sqrt{V\nu(x,t)}\dot W(x,t).
\end{equation}
As mentioned above, we refer to SPDE (\ref{SPDE}) as the Stochastic Asexual Gaussian allelic model with Abundance dynamics (abbreviated SAGA). Comparing equation (\ref{SPDE}) to equation (3.5) of Li [-@zeng1998absolute], our $m$ and $V$ correspond to Li's $b$ and $c$ respectively. It is important to note that, under the assumptions made in Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] and Li [-@zeng1998absolute], equation (\ref{SPDE}) is only formal when $m(h,x)$ is bounded across all combinations of $h\geq0$ and $x\in\mathbb{R}$. However, recalling our condition $m(h,x)\leq r\in\mathbb{R}$, the growth rates we consider are only bounded above. Yet, in the proof of the construction of the interacting superprocess as the limit of rescaled branching diffusions, Méléard and Roelly [-@meleard1992interacting; -@Mlard1993] assumed $m(h,x)$ to be bounded to guarantee the total mass process will have finite mean and variance, for finite $t\geq0$. This allowed the authors to employ a tightness criterion for sequences of measures and show the rescaled processes converge to a superprocess with finite total mass. Li's [-@zeng1998absolute] result builds directly off of Méléard and Roelly's construction, inheriting the assumption of boundedness for $m(h,x)$. However, in Li [-@zeng1998absolute], the sufficiency of $m(h,x)$ being bounded above is even more clear since Li works explicitly with a common upperbound for both $m(h,x)$ and $V$. Hence, one can repeat the necessary proofs replacing the assumption that $m(h,x)$ is bounded with the assumption that $m(h,x)$ is merely bounded above to derive the same results.

@@

Our treatment of the relationship between breeding values and expressed traits follows classical quantitative genetic assumptions such as those used in the seminal paper by Lande [-@Lande1975] to investigate the maintenance of genetic variation. In particular, we assume that the expressed trait for any given individual is independent of environmental conditions and normally distributed around their breeding value with variance $\eta$. Hence, $\sigma^2=G+\eta$. In the case that all of the effects of alleles on an expressed trait are additive, $\eta$ is known as the _variance of environmental deviation_ [@Lande1975; @michaellynch1998]. For a given breeding value, we denote the probability density of a randomly drawn expressed trait by $\psi(x,g)$. Hence,
\begin{equation}
\psi(x,g)=\frac{1}{\sqrt{2\pi\eta}}\exp\left(-\frac{(x-g)^2}{2\eta}\right).
\end{equation}
To include this relationship in our framework, we write $\rho(g,t)$ as the abundance density of breeding values at time $t$ so that $\int_{-\infty}^{+\infty}\rho(g,t)dg=\int_{-\infty}^{+\infty}\nu(x,t)dx=N(t)$. We switch our focus from directly modelling the evolution of $\nu(x,t)$ to modelling the evolution of $\rho(g,t)$. Once $\rho(g,t)$ is determined, we can compute $\nu(x,t)$ via
\begin{equation}\label{nu_from_rho}
\nu(x,t)=\int_{-\infty}^{+\infty}\rho(g,t)\psi(x,g)dg.
\end{equation}
However, since selection acts on expressed phenotypes, we use the assumed relationship between breeding values and expressed traits to calculate the fitness of breeding values. To motivate the approach taken, consider the problem of inferring the breeding value of an individual given its expressed trait $x$. Denote $\mathfrak{g}$ a random variable representing the unkown breeding value. Under this model of inheritance we know $x$ is a random sample from a normal distribution with mean $\mathfrak{g}$ and variance $\eta$. Maximizing likelihood suggests $x$ is our best guess for $\mathfrak{g}$, but the actual value of $\mathfrak{g}$ is normally distributed around $x$ with the variance $\eta$. Hence, for fixed $x$, we obtain $\psi(x,g)$ as the probability density of $\mathfrak{g}$. Thus, the mean fitness of a breeding value $g$ across all individuals carrying $g$ can be written as
\begin{equation}\label{m_rel}
m^*(\rho,g)=\int_{-\infty}^{+\infty}m(\nu,x)\psi(x,g)dx.
\end{equation}
This is similar to the approach taken by Kimura and Crow [-@Kimura1978] to calculate the overall effects of selection for expressed characters onto the changes in the distribution of alleles encoding those characters. However, instead of focusing on the frequencies of alleles at particular loci, our results focus on the densities of breeding values. With the relationship between $m(\nu,x)$ and $m^*(\rho,g)$ established, we define the evolution of $\rho(g,t)$ by the SPDE
\begin{equation}\label{rho_SPDE}
\dot\rho(g,t)=m^*(\rho,g)\rho(g,t)+\frac{\mu}{2}\frac{\partial^2}{\partial^2 g}\rho(g,t)+\sqrt{V\rho(g,t)}\dot W(g,t).
\end{equation}
Equation (\ref{rho_SPDE}) is a stochastic generalization of the deterministic PDE (\ref{eq1}) from \S\ref{deterministic}, but describes the evolution of the distribution of breeding values instead of expressed characters. However, whether modelling expressed characters or breeding values, we refer to SPDE of the form (\ref{SPDE}) as Stochastic Asexual Gaussian allelic models with Abundance dynamics (abbreviated SAGA). By replacing $\psi(x,g)$ with different probability densities produces different developmental models. This includes densities that may depend on environmental variables and hence model phenotypic plasticy. To leave room for the naming of other models where $\psi(x,g)$ is Gaussian, we refer to the model of development we have chosen as the _centered_ Gaussian developmental model. In \S\ref{wnc} we review the origins of this equation and provide some theory to help make sense of it, particularly the term $\dot W$.

@@

```{r, message=FALSE, echo=F, fig.align='center', out.width='100%', fig.cap='\\label{temporal}Temporal dynamics of mean trait (top), additive genetic variance (middle) and abundance (bottom) for the scenario of weak competition (left) and strong competition (right).'}
knitr::include_graphics('community_dynamics.jpg')
```

@@

We now make use of the expressions derived for competition coeffecients and selection gradients to investigate their relationship. As a first pass, let us assume all model parameters are equivalent across species and that each species has the same abundance and trait variance. Let us further assume that richness $S$ is large and the distribution of mean trait values is normal with mean $\bar{\bar x}$, variance $V_{\bar X}$ and density $f_{\bar X}$. Such assumptions are typical when deriving analytical results in the field of theoretical coevolutionary community ecology [@Nuismer2012; @Nuismer2018]. If $\bar{\bar x}$ is near $\theta$ and $V_{\bar X}$ is much smaller than $|2R/a-G-\eta|$, then we may approximate $r_i$ with

@@

We have introduced an approach to derive models of evolutionary ecology using the calculus of white noise, developed SAGA, a novel model of phenotypic evolution that accounts for demographic stochasticity, demonstrated our approach by deriving a model of diffuse coevolution from classical competition theory, and used this model to investigate the relationship between selection gradients and competition coefficients. This work demonstrates that connecting contemporary theoretical approaches to evolutionary ecology with some fundamental results in the theory of measure-valued Markov processes allows for the development of a rigorous, yet flexible approach to synthesizing the dynamics of abundance and distributions of quantitative characters. Although this approach enjoys these important merits, there remains biological details and their associated technical challenges that need to be confronted for gaining a more thorough and accurate understanding of ecological communities. We touch on just four of them here and provide suggestions for future research to approach these challenges.

__Limitations of diffusion limits and validating coexistence theory__

As noted by Feller [-@feller1951], diffusion limits provide reasonable approximations for large populations, but relatively small populations require discrete models. Hence, as a diffusion limit, SPDE (\ref{SPDE}) restricts parameter values to regions that maintain large population sizes. Although this suggests an important restriction on any model developed under this framework by implying populations are not at risk of extinction, the SDE describing abundance dynamics technically permits extinction. However, for small abundances, pathology emerges in the evolution of trait means and variances. In particular, stochastic components of the SDE describing the evolution of $\bar x$ and $\sigma^2$ diverge towards infinite values as $N\to0$. Hence, studies of evolutionary rescue and colonization-extinction dynamics that incorporate phenotypic evolution should be pursued via a different approach. A natural alternative can be developed utilizing the underlying BBM that SPDE (\ref{SPDE}) is a diffusion limit of (see section \ref{stochastic}). This process explicitly tracks individuals throughout their life-history and captures reproduction as branching events. Hence, BBM processes model population size as a non-negative integer instead of a continuously varying number. In particular, the pathological behavior described for the diffusion limit does not occur for BBM as population size goes to zero. Relating this concern to our comment in \S\ref{ecoevo} on using our model of diffuse coevolution to extend modern coexistence theory [sensu @Chesson2000] in a way that accounts for phenotypic evolution and abundance dynamics, it seems necessary to determine whether or not results pertaining to the coexistence of competing populations based on deterministic models that ignore phenotypic evolution hold using stochastic individual-based simulations. To stay within the framework outlined in this paper and thereby produce comparable results, these simulations should be extensions of BBM that account for the effects of competition.

@@

Here we set out to show, given the initial conditions $\nu(\cdot,0)\in L^1(\mathbb{R})$ such that $\bar x(0)\in\mathbb{R}$, $\sigma^2(0)\in[0,+\infty)$ and $N(0)\in(0,+\infty)$, and given the growth rate
\begin{equation}\label{m_log_stab}
m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_\mathbb{R}\nu(y,t)dy
\end{equation}
such that $\theta\in\mathbb{R}$, $a,c,\mu>0$ and $r>\tfrac{1}{2}\sqrt{\mu a}$, we have the following stable equilibrium $N_\infty=\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})$, $\bar x_\infty=\lim_{t\to\infty}\bar x(t)=\theta$ and $\sigma^2_\infty=\lim_{t\to\infty}\sigma^2(t)=\sqrt{\tfrac{\mu}{a}}$.
